\textbf{SPECFEM3D}\\ \textbf{User Manual}

### 2010-11-22 doc/USER_MANUAL/manual_SPECFEM3D.tex
updated all the headers
final modifications to the manual; changed "Harvard" to "Princeton" on the cover; added a sentence about conforming mesh doublings

%% LyX 1.5.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[oneside,english,onecolumn]{book}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{float}
\usepackage{textcomp}
\usepackage{amsmath}

@@

The software package SPECFEM3D simulates seismic
wave propagation at the local or regional scale
based upon the spectral-element method (SEM).
The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.
Note that in most geological models in the context of seismic wave propagation studies
(except for fault dynamic rupture studies)
a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient. For a detailed introduction to the SEM as applied to regional
seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08} and
in particular \citet{KoLiTrSuStSh04}.
A detailed theoretical analysis of the dispersion
and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07} and \citet{SeOl07}.

Effects due to lateral variations in compressional-wave speed, shear-wave
speed, density, a 3D crustal model, topography and bathymetry are
included. The package can accommodate
full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
kernel simulations are included \citep{LiTr06,TrKoLi08,FiIgBuKe09,ViOp09}.

All SPECFEM3D software is written in Fortran90 with full portability
in mind, and conforms strictly to the Fortran95 standard. It uses
no obsolete or obsolescent features of Fortran77. The package uses
parallel programming based upon the Message Passing Interface (MPI)
\citep{GrLuSk94,Pac97}.

SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
conference in Phoenix, Arizona (USA) (see \cite{KoTsChTr03}
and \url{www.sc-conference.org/sc2003/nr_finalaward.html}).
It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.
It also won the BULL Joseph Fourier supercomputing award in 2010.

@@

\item [Forward simulations] ~\\
\cite{TrKoLi08,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}

\item [Adjoint simulations] ~\\
\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoLi08,TrKoHjLiZhPeBoMcFrTrHu10}

@@

The corresponding Bib\TeX{} entries may be found
in file \texttt{USER\_MANUAL/bibliography.bib} or in comments at the
beginning of file \texttt{specfem3D.f90}.

@@

$\copyright$ Princeton University and University of Pau / CNRS / INRIA, February 2010

>>>>

%% LyX 1.5.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[oneside,english,onecolumn]{book}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{geometry}
\geometry{letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{float}
\usepackage{textcomp}
\usepackage{amsmath}

@@

The software package SPECFEM3D simulates seismic
wave propagation at the local or regional scale
based upon the spectral-element method (SEM).
The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06,WiStBuGh10}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.
Note that in most geological models in the context of seismic wave propagation studies
(except for fault dynamic rupture studies)
a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient;
conforming mesh doubling bricks can efficiently handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04}.\\

For a detailed introduction to the SEM as applied to regional
seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08} and
in particular \citet{KoLiTrSuStSh04}.
A detailed theoretical analysis of the dispersion
and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07} and \citet{SeOl07}.\\

Effects due to lateral variations in compressional-wave speed, shear-wave
speed, density, a 3D crustal model, topography and bathymetry are
included. The package can accommodate
full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
kernel simulations are included \citep{LiTr06,TrKoLi08,FiIgBuKe09,ViOp09}.\\

All SPECFEM3D software is written in Fortran90 with full portability
in mind, and conforms strictly to the Fortran95 standard. It uses
no obsolete or obsolescent features of Fortran77. The package uses
parallel programming based upon the Message Passing Interface (MPI)
\citep{GrLuSk94,Pac97}.\\

SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
conference in Phoenix, Arizona (USA) (see \cite{KoTsChTr03}
and \url{www.sc-conference.org/sc2003/nr_finalaward.html}).
It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.
It also won the BULL Joseph Fourier supercomputing award in 2010.\\

@@

\item [Numerical simulations in general] ~\\
Forward simulations are described in detail in \cite{TrKoLi08,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}. Additional aspects dealing with adjoint simulations are described in
\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10}.

@@

The corresponding Bib\TeX{} entries may be found
in file \texttt{doc/USER\_MANUAL/bibliography.bib}.

@@

$\copyright$ Princeton University and University of Pau / CNRS / INRIA, November 2010
